QPLIB: A Library of Quadratic Programming Instances

### 2018-02-12 QPLIB-2.2.tex
problem types should be formatted in italics
Update QPLIB-2.2.tex

We now give a general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, we will assume minimization problems for the remaining of this section.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, D*B problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, {\textstyle \half} Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here. Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context. At the other end of the spectrum, in general QP is a hard problem. Actually, \textit{LIQ}---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 {\textstyle \half} x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.

@@

This description is deliberately coarse; each of these classes can be subdivided into several others on the grounds of more detailed information about structures present in their constraints/objective function. These can have a significant algorithmic impact, and therefore can be of interest to researchers. Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11}\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}, or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}\todo{SV: 5 Frangioni et.al citations for SOC?}. It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP. For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this section.

>>>>

We now give a general discussion about the different problem classes that our proposed taxonomy defines.
For simplicity, we will assume minimization problems for the remaining of this section.
Some problem classes are actually ``too simple'' to make sense in our context. For instance, \textit{D*B} problems have only diagonal quadratic (hence separable) objective function and bound constraints; as such, they read
\[
 \textstyle
 \min \; \Big\{ \,
 \sum_{j \in \Mcn} \big( \, {\textstyle \half} Q^0_j x_j^2 + b^0_j x_j \, \big) \;:\;
 l_j \leq x_j \leq u_j \quad j \in \Mcn \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcz \, \Big\}
 \;\; .
\]
Hence, their solution only requires the independent minimization of a convex quadratic univariate function in each single variable $x_j$ over a box constraint and possibly integrality requirements, which can be attained trivially in $O(1)$ operations (per variable) by closed-form formul{\ae}, projection and rounding arguments. \emph{A fortiori}, the even simpler cases \textit{L*B}, \textit{D*N} and \textit{L*N} (the latter obviously unbounded unless $b^0 = 0$) will not be discussed here. Similarly, \textit{CCN} are immediately solved by linear algebra techniques, and therefore are of no interest in this context. At the other end of the spectrum, in general QP is a hard problem. Actually, \textit{LIQ}---linear objective function and quadratic constraints in integer variables with no finite bounds, i.e.
\[
 \textstyle
 \min \; \Big\{ \, b^0 x \;:\;
 {\textstyle \half} x^\top Q^i x + b^i x \leq c^i \quad i \in \Mcm \;\;,\;\;
 x_j \in \mathbb{Z} \quad j \in \Mcn \, \Big\}
 \;\; ,
\]
is not only $\mathcal{NP}$-hard, but downright undecidable \cite{jeroslow}. Hence so are the ``harder'' \{\textit{C},\textit{Q}\}\textit{IQ}.

@@

This description is deliberately coarse; each of these classes can be subdivided into several others on the grounds of more detailed information about structures present in their constraints/objective function. These can have a significant algorithmic impact, and therefore can be of interest to researchers. Common structures are, e.g., network flows \cite{FrFG16,FrGaSc14,FrGaSt16a,FGGP11,tadayon-smith:2013} or knapsack-type linear constraints \cite{FrFG16,FrGo13b,FGGP11}, and semi-continuous variables \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe06a,FrGe07a,FrGe09a,FGGP11},
%\todo{SV: do we really need 7 Frangioni et.al citations for semi-continuous variables?}
or the fact that a nonconvex quadratic objective function/constraint can be reformulated as a second-order cone (hence, convex) one \cite{FrFG16,FrGaSc14,FrGaSt16a,FrGe09a,FGGP11}.
%\todo{SV: 5 Frangioni et.al citations for SOC?}. 
It would be very hard to collect a comprehensive list of all types of structures that might be of interest to any individual researcher, since these are as varied as the different possible approaches for specialized sub-classes of QP. For this reason we do not attempt such a more refined classification, and limit ourselves to the coarser one described in this section.
